% A pair of parragraphs introducing this manual.

This document explains how to use Scoja.
If you know how to write Syslog-ng configuration files, and you are
not interesting in learning Scoja native configuration language, then
you only need to read Section~\label{conf}, where {\tt scoja} program
options are explained.
The same is true for tradicional syslog configuration files,
because Scoja understands both syslog and Syslog-ng configuration
languages.

But to get the best of Scoja, it is necessary to use Scoja native
configuration lenguage (SCL).
SCL main concepts are those of Syslog-ng:
sources, filters and targets linked together.
SCL has other concepts,
like actions, function, an event context, template user variables,
etc.
With all these elements, a \emph{logging net} is defined that process
events from sources to targets, discarding unimportant ones.
But the distintive caracteristic of SCL is that \emph{it is Python},
with a few definitons added to build the logging net.
This choose is a clear winner because all the power of Python can be
used to organize and simplify configuration:
  variables, funcion definition, arrays, loops, conditionals, etc.
But it is not necessary to learn Python to write SCL:
if you define and combine elements following instructions in this
manual, you will get fully functional configurations files that use
all of Scoja logging capabilities.
What you loose not knowing Python is a shorter, best organized
configuration file.

If you like Scoja, and are planning to use it with complex
configuration, we recommend you to know a bit of Python,
just how to define functions, and how arrays work.
Try~\cite{Diving} or~\cite{Python}.
